---
title: "R Notebook"
output: html_notebook
author: Lynn Marciano
---

## Simple EDA

```{r}
library(tidyverse)
library(ggplot2)
library(patchwork)
library(sandwich)
library(lmtest)
```

```{r}
filtered_listings_data <- read.csv("filtered_listings_data.csv")
```


```{r}
#subtracting availability from 60 to see how many bookings there are within the next 60 day time span
listings_bookings <- filtered_listings_data %>%
  mutate(
    bookings_next_60 = 60 - availability_60
  )
glimpse(listings_bookings)
```


```{r}
listings_bookings_histogram <- listings_bookings %>%
  ggplot()+
  aes(x = bookings_next_60)+
  geom_histogram(bins=30)+
  labs(
    title = "Total Number of Bookings",
    x = "Bookings",
    y = "Count"
  )

listings_bookings_histogram

```

> Total Number of bookings per the next 60 days is definitely modal.

```{r}
#looking at the distribution of ratings

listings_rating_histogram <- listings_bookings %>%
  ggplot()+
  aes(x = review_scores_rating)+
  geom_histogram()+
  labs(
    title = "Listing Ratings",
    x = "Rating",
    y = "Count"
  )

listings_rating_histogram
```

> Ratings are left skewed, a negative skewness. Mode at 1.00 which is the highest rating.


## Ratings and Bookings by themselves

```{r}
scatter_listings_rating_bookings <- listings_bookings %>%
ggplot()+
aes(y = review_scores_rating, x = bookings_next_60)+
geom_point()+
labs(
title = "Relationship of Bookings and Rating",
x = "Bookings",
y = "Ratings"
)

scatter_listings_rating_bookings
```

> There are a few concerns with our research question as there may not be a defined relationshiping with ratings and bookings..


## Does the total number of days the listing is active affect the number of bookings and ratings?

It is under assumption that the listings that are most recently active will have a higher rating. Newer listings may also have more availability; or less bookings, then established listings.

Since there is no active_since column, I'll use host_since. However, this could be misleading since vacation rental companies may be a host for awhile, but have just recently listed a property. On another note, we could analyze if hosts that have been hosts longer could potentially affect the number of bookings a listing may have.

```{r}
#We have the last_scraped date and the host_since date. We will subtract the host_since date from last_scraped to get a rough estimate as to how long the listing has been active; or at the very least how long the host for the listing has been a host. 

listings_bookings <- listings_bookings %>%
  mutate(
    host_active_days = as.numeric(
      difftime(as.Date(as.character(last_scraped), format="%m/%d/%y"),
                  as.Date(as.character(host_since), format="%m/%d/%y"),units="days"))
  )

summary(listings_bookings$host_active_days)
max(listings_bookings$host_active_days)
min(listings_bookings$host_active_days)
```

```{r}
listings_rating_host_scatterplot <- listings_bookings %>%
  ggplot()+
  aes(x = host_active_days, y= review_scores_rating)+
  geom_point()+
  labs(
    title = "Listing Ratings based on Host Active Days",
    x = "Host Active Days",
    y = "Rating"
  )

listings_rating_host_scatterplot
```
> I am not sure if we can see a distinguishable correlation here. However, it is a little bit apparent that the hosts who have been listing the longest are not likely to have a lower rating. In fact, most of the long running hosts have ratings no less than .7


```{r}
listings_bookings_host_scatterplot <- listings_bookings %>%
  ggplot()+
  aes(x = host_active_days, y= bookings_next_60)+
  geom_point()+
  labs(
    title = "Listing Bookings based on Host Active Days",
    x = "Host Active Days",
    y = "Bookings"
  )

listings_bookings_host_scatterplot
```

> There is absolutely no distinction between number of bookings based on the duration a host has been active.

---------------------------------------------------------------------------

# Model Building

### I've decided to approach the model building by backwards selection.

I've removed all the bookings variables since we are only interested in bookings_next_60. 

I've also removed other variables that would not be active coefficients to explore.
The variables include: id, host_id, last_scraped, neighbourhood_group_cleansed, calendar_last_scraped, first_review, last_review.

I removed bathrooms_text because I am not sure how we would handle and distinguish between shared or private baths.

I removed neighbourhood_cleansed because there were 259 different categories and our data is based on LA as a whole.

After analyzing host_since and seeing very little correlation, that variable was removed.

```{r}
listings_bookings_final = subset(listings_bookings, select=-c(
  availability_30,
  availability_60,
  availability_90,
  availability_365,
  id, 
  host_id,
  neighbourhood_cleansed,
  last_scraped, 
  neighbourhood_group_cleansed,
  calendar_last_scraped,
  first_review,
  last_review,
  bathrooms_text,
  host_since,
  calculated_host_listings_count,
  calculated_host_listings_count_entire_homes,
  calculated_host_listings_count_private_rooms ,
  calculated_host_listings_count_shared_rooms))
glimpse(listings_bookings_final)
```

We're left with room_type and property_type as the only categorical data. I changed this to a numerical categorical for the regression model.

```{r}
listings_bookings_final$room_type <- as.numeric(factor(listings_bookings_final$room_type))

listings_bookings_final$property_type<- as.numeric(factor(listings_bookings_final$property_type))
```

```{r}
glimpse(listings_bookings_final)
```
The new numerical categories are:

- 1 - Entire home/apt
- 2 - Hotel room
- 3 - Private room
- 4 - Shared room


```{r}
#Build a model with all the coefficients
model <- lm(bookings_next_60 ~ ., data = listings_bookings_final)
coeftest(model, vcov= vcovHC(model))
```
Based on our t-test of all our coefficients, we can see the following variables are significant:

- host_is_superhost
- room_type
- property_type
- accommodates                
- price
- minimum_nights
- has_availability (we will most likely not include this)
- number_of_reviews_l30d
- review_scores_cleanliness
- review_scores_value
- host_active_days
- instant_bookable
- number_of_reviews
 
Interestingly enough, the variable that we are trying to measure has an effect on bookings is not significant at all, with a p-value of .75

We can also observe that the maximum and minimum amount of nights does not affect the booking.

Surprisingly, beds and bedrooms are not significant but accommodates is. This means we could potentially have a one bedroom apartment that accommodates 10 and it may be more likely to be booked.

*We may want to consider rephrasing our question to **What causes more bookings?**\ *

### Reducing the model to the significant coefficients
```{r}
model1 <- lm(bookings_next_60 ~ host_is_superhost + room_type + property_type + accommodates + price + minimum_nights + number_of_reviews_l30d + review_scores_cleanliness + review_scores_value + host_active_days + instant_bookable + number_of_reviews + host_listings_count + host_identity_verified, data = listings_bookings_final)

coeftest(model1, vcov= vcovHC(model1))
```
### Checking different fits

```{r}
model2 <- lm(bookings_next_60 ~ host_is_superhost + room_type + property_type + accommodates + price + minimum_nights + number_of_reviews_l30d + review_scores_cleanliness + review_scores_value + host_active_days + instant_bookable + number_of_reviews , data = listings_bookings_final)

coeftest(model2, vcov= vcovHC(model2))
```
> A good review_scores_value is associated with 300 more bookings.
> A superhost is associated with 1.2 more bookings
> A listing that is instantly bookable is associated with is 45% more likely to get a booking.

> property_type, room_type, accomodates, price, minimum_nights, number_of_reviews_l30d, and review_scores_cleanliness are all associated with more bookings but at a diminishing rate


```{r}
# adding back in review_scores_rating
model3 <- lm(bookings_next_60 ~ review_scores_rating + host_is_superhost + room_type + property_type + accommodates + price + minimum_nights + number_of_reviews_l30d + review_scores_cleanliness + review_scores_value + host_active_days + instant_bookable + number_of_reviews , data = listings_bookings_final)

coeftest(model, vcov= vcovHC(model3))
```

We will assume a difference in slope between a superhost and regular host; or a difference in the marginal returns for ratings with superhosts and hosts

```{r}
model4 <- lm(bookings_next_60 ~ review_scores_rating + host_is_superhost +
               host_is_superhost*review_scores_rating + room_type +
               property_type + accommodates + price + minimum_nights +
               number_of_reviews_l30d + review_scores_cleanliness +
               review_scores_value + host_active_days + instant_bookable +
               number_of_reviews , data = listings_bookings_final)

coeftest(model4, vcov= vcovHC(model4))
```

> review_scores_rating.host_is_superhost coefficient is significant. However, review_scores_rating is still insignificant.

 wonder if review_scores_rating is not significant because it is on a scale of 1.
I am going to multiply the review_scores_rating by 5 since on airbnb the scale is 5 stars.

```{r}
listings_bookings_final <- listings_bookings_final %>%
  mutate(rating_amped = review_scores_rating*5)

```

```{r}
model5 <- lm(bookings_next_60 ~ rating_amped + host_is_superhost + room_type +
               property_type + accommodates + price + minimum_nights +
               number_of_reviews_l30d + review_scores_cleanliness +
               review_scores_value + host_active_days + instant_bookable +
               number_of_reviews , data = listings_bookings_final)

coeftest(model5, vcov= vcovHC(model5))
```

We saw earlier that rating was not linear. I'll apply a log transform to the rating to try to make it more normal.
The residuals have a skewed distribution, so we will apply this transform to obtain residuals that are a little bit more symmetrically distributed and to linearize the relationship of ratings and bookings since that is our original research question.

```{r}
model6 <- lm(bookings_next_60 ~ log(review_scores_rating) + host_is_superhost + 
               room_type +
               property_type + accommodates + price + minimum_nights +
               number_of_reviews_l30d + review_scores_cleanliness +
               review_scores_value + host_active_days + instant_bookable +
               number_of_reviews , data = listings_bookings_final)

coeftest(model6, vcov= vcovHC(model6))
```
> A log transform on review_scores_rating makes it a bit more significant but at a diminishing rate.
> We see that review_scores_value is still very significant, with 3.44 more bookings 
> Instant bookable doesnt seem as significant anymore when we linearized rating

```{r}
model7 <- lm(bookings_next_60 ~ log(review_scores_rating) + host_is_superhost +
               room_type + property_type + accommodates + price + minimum_nights +
               number_of_reviews_l30d + review_scores_cleanliness +
               review_scores_value + host_active_days +
               number_of_reviews , data = listings_bookings_final)

coeftest(model7, vcov= vcovHC(model7))
```
I wonder if having another review_scores affects the rating.
I also remove the number_of_reviews.

```{r}
model8 <- lm(bookings_next_60 ~ log(review_scores_rating) + host_is_superhost +
               room_type + property_type + accommodates + price + minimum_nights +
               number_of_reviews_l30d + 
               review_scores_value + host_active_days,
             data = listings_bookings_final)

coeftest(model8, vcov= vcovHC(model8))
```
Removing review_scores_value

```{r}
model9 <- lm(bookings_next_60 ~ log(review_scores_rating) + host_is_superhost +
               room_type + property_type + accommodates + price + minimum_nights +
               number_of_reviews_l30d +  host_active_days,
             data = listings_bookings_final)

coeftest(model9, vcov= vcovHC(model9))
```

Back to where we started. It seems model7 is the best.